### 2018-04-12 paper/algorithm.tex
finished route_success_rate plot
described my new idea

removed big'o'file

\NewCryptoEntity{\Alice}{A}
\NewCryptoEntity{\Bob}{B}
\NewCryptoEntity{\Carol}{C}
\NewCryptoEntity{\David}{D}

We consider several participants: Alice, Bob, Carol and David.
Each of them has a set of devices, \(D_i\) (where \(i\in \{\Alice, \Bob, \Carol, 
  \David\}\)), hereafter called their device swarm.
Our system is composted of two network overlays: one protocol is run between the 
devices in \(D_i\) (the Device Swarm Protocol) to gain knowledge on the 
swarm, and another run between different swarms (the Inter-Swarm Protocol) to 
keep global knowledge.
Finally, our file-sharing protocol uses the local and global 
knowledges to coordinate file sharing between devices in different swarms.

\subsection{The Device Swarm Protocol}%
\label{DeviceSwarmProtocol}

A user's devices, \(D_i\), use the Device Swarm Protocol to coordinate 
themselves.
More specifically, the devices use this protocol to keep the swarm state up to 
date.
The swarm state contains:
\begin{itemize}
  \item A knowledge vector, \(A_i\), encoding a Markov chain to predict online 
    and offline patterns for the various devices to predict when and for how 
    long each device might be available before disappearing again;
  \item A table, \(F_i = N_i\times \powerset(D_i)\), mapping a user's file names 
    (\(N_i\)) to its devices (\(D_i\)) that store a copy of that file.
\end{itemize}

\commentDaniel{What should we reuse from the session-handoff paper?}

\subsection{The Inter-Swarm Protocol}%
\label{InterSwarmProtocol}

All swarms participate in the Inter-Swarm Protocol to coordinate the global 
knowledge needed by the swarms to interact.
More specifically, the global knowledge contains:
\begin{itemize}
  \item A table, \(R\), of devices available for routing;
  \item A key--value store, \(S = \{0, 1\}^\lambda\times \{0, 1\}^*\), mapping 
    fixed-length strings of length \(\lambda\) to arbitrary-length strings.
\end{itemize}

\subsection{File-sharing protocol}%
\label{FileSharingProtocol}

\begin{figure}[t]
\centering
\includegraphics[width=0.8\columnwidth]{figures/schema.pdf}
\caption{\label{fig:pipeline}Pipeline of the steps undergone by Alice when he shares a file from his laptop L with Bob.}
\end{figure}

\Cref{fig:pipeline} depicts the process of a file exchange from Alice to Bob.
The file $f$ is initially located on Alice's laptop $L$, which is potentially offline for now.
Firstly, Alice and Bob need to exchange encryption keys and the key $ID_f$ 
referencing the file on the \ac{DHT}.
They do so \emph{out-of-band}, that is using another channel than the one 
presented here (\eg bluetooth, email, \etc).

Then, Alice queries the DHT for devices that will most likely be online at the same time as his laptop $L$.
To this end, Alice uses his knowledge on his devices connection patterns $A_i$, along with the DHT's public information on devices connections.
Once Alice and the DHT agreed upon a set of peers $(M_1, M_2, M_3)$, their address is put in the DHT at the key $ID_f$ referencing the file of interest.

At the third step, Alice's laptop $L$ becomes available. 
Informed of the ongoing file exchange by Alice's Device Swarm Protocol, $L$ fetches the addresses of $M_1, M_2$ and $M_3$ on the DHT using the key $ID_f$.
These peers will serve as middle-men: $L$ sends them redundant chunks of $f$ through an onion route (thus remaining anonymous), for Bob to fetch them later.

In the last forth step, Bob fetches the file $f$ using any of her devices.
To do so, she uses the key $ID_f$ to learn the addresses of $M_1, M_2$ and $M_3$.
Then, she fetches parts of $f$ from the online middle-men through an onion route.
Given that $L$ sent redundant chunks to the middle-men, Bob will not need every middle-man to be connected to fetch the whole file.
Once Bob has completed the download, the key $ID_f$ is flushed on the DHT.

% When Alice and Bob want to exchange a file owned by Alice, they first need to exchange their

% When Alice wants to transfer a file to Bob, she will find which devices have the 
% file (from \(F_i\)), then predict when any of those devices will be available 
% (from \(A_i\)).
% Then she will construct a probabilistic onion route with her devices as the 
% destinations.
% She will encrypt this route for Bob and transfer it to him, who now can use the 
% onion route to fetch the file.

>>>>

% \NewCryptoEntity{\Alice}{A}
% \NewCryptoEntity{\Bob}{B}
% \NewCryptoEntity{\Carol}{C}
% % Arg, already used \mathcal{D} for the devices set
% \NewCryptoEntity{\David}{D}

\newcommand{\Alice}{\ensuremath{\mathscr{A}}}
\newcommand{\Bob}{\ensuremath{\mathscr{B}}}

\newcommand{\localoverlay}{Squad Overlay}
\newcommand{\globaloverlay}{Global Overlay}

A set of users $\mathcal{U}$ participate in the \name network.
Each user $u \in \mathcal{U}$ has a set $\mathcal{D}_u$ of devices that contribute to the system when online.
Alice and Bob use \name to exchange a file secretly: no participant will learn who sends and receives it, nor its content.

On one hand, a user's devices must exchange information among themselves to learn the user's behavior \commentAL{Can we skip files mapping, pretty please?} and map files to devices.
We call this part of the protocol the \localoverlay, and present it in section~\ref{local overlay}. 
On the other hand, user's collectively upload the list of their connected devices and associated future predictions to a Distributed Hash Table (DHT): it serves as a registry of online peers (including connection probabilities for the next turn), and for the routing of files.
We describe this \globaloverlay in section~\ref{global overlay}. 
File exchange between two users is enabled by the two previous protocols, and is covered in section~\ref{file exchange}

\subsection{The \localoverlay}%
\label{local overlay}

A user's devices, \(D_i\), use the \localoverlay to:
\begin{enumerate}
	\item Share information about the user's behavior.
	Devices use it to build a behavioral model and infer their future connection probability;
	\item Share files location \commentAL{Maybe...}
\end{enumerate}

\subsubsection{Sharing the user's behavior} % (fold)
\label{ssub:sharing_the_user_s_behavior}

Each device initially only knows when it is connected. 
To make a model of the user's behavior, they need to know the whole observation sequence $O$: which device is online at each time step.
To do so, they employ a probabilistic dissemination protocol, based on previous work from our team \cite{luxey:hal-01704172}.
This algorithm, called \textsc{Sprinkler}, allows devices to gossip any new connection information to random peers, ensuring that all devices know the full observation sequence with a very high probability.
\textsc{Sprinkler} has shown resilient to devices churn \cite{luxey:cascade}, and is thus perfectly suited for our purpose, where a user's devices often disconnect and reconnect.

\commentAL{Then we \textbf{maybe} do predictions as \textbf{maybe} shown in section~\ref{sec:user_model}.}

% subsubsection sharing_the_user_s_behavior (end)

\subsubsection{Sharing files location} % (fold)
\label{ssub:sharing_files_location}

We do stuff and things work.

% subsubsection sharing_files_location (end)

\subsection{The \globaloverlay}%
\label{global overlay}

% All swarms participate in the Inter-Swarm Protocol to coordinate the global 
% knowledge needed by the swarms to interact.
% More specifically, the global knowledge contains:
% \begin{itemize}
%   \item A table, \(R\), of devices available for routing;
%   \item A key--value store, \(S = \{0, 1\}^\lambda\times \{0, 1\}^*\), mapping 
%     fixed-length strings of length \(\lambda\) to arbitrary-length strings.
% \end{itemize}

Every device in the system participates in a Distributed Hash Table (DHT) that is used for two purposes:
\begin{enumerate}
	\item Providing routing information to nodes during file exchange. This is discussed in-depth in section~\ref{file exchange};
	\item Listing online devices and their probability of being online next time step.
\end{enumerate}

Every time a device is online, it sends the following information to the DHT:
\begin{itemize}
	\item the time \commentAL{step?};
	\item its address;
	\item its public key;
	\item the probability that it will still be connected next turn \commentAL{no interest in t+2 and more.}
\end{itemize}

\commentAL{Cite DHT articles and emphasize their security guarantees.}

\subsection{The file exchange}%
\label{file exchange}

We consider two users, Alice and Bob, such that Alice wants to send a file to Bob secretly.
They want to keep the file content secret to anybody but them; \name's users should not know who sent a file to whom; and Alice and Bob want to hide their address from one another.

In other words, \name must guarantee \emph{sender-receiver anonymity}.
To achieve this, communications go through onion routes, using rendez-vous points.
Our approach is tightly based on the Tor seminal paper \cite{Tor}.

\subsubsection{File exchange process} % (fold)
\label{ssub:file_exchange_process}

Firstly, Alice and Bob must exchange information out-of-band: an encryption key $ek$ to cipher the file, and an ID $k$ in the DHT pointing to information on how to reach Alice.

The device that contains Alice's file \commentAL{I don't like that dependence on the file's location} creates routes to several Introductory Points (IP), chosen at random among connected peers.
Alice uploads the addresses of her IP to the DHT at the position $k$.
One of Bob's devices creates routes to several Rendez-Vous points (RV), randomly picked among peers that will be connected in the next time step.
Bob fetches the addresses of Alice's IP from the DHT.
Then, he sends a packet containing the addresses of its chosen RV to Alice's IPs.
The IP nodes forward the packet to Alice's device. 
Alice chunks her file into pieces of a predefined size, encrypted using $ek$.

Starting from the next round, Alice will send all chunks of her file to Bob's RV, that will forward it to Bob.
Once the file is received, everybody is happy and all routes are killed.

% subsubsection file_exchange_process (end)

\subsubsection{Onion routes in face of churn} % (fold)
\label{ssub:onion_routes_in_face_of_churn}

An onion route between A and B consists of several servers (usually three) that serve as relays, say $r_1, r_2$ and $r_3$. 
The route will have a fixed order: $A \rightarrow r_1 \rightarrow r_2 \rightarrow r_3 \rightarrow B$.
Every time data travels through the onion route, it is encrypted---along with its header---with symmetric keys, such that no-one except the sender and the receiver can read its content.
A node only knows its direct predecessor (that sent the packet) and its direct successor (whose address is the only one readable by it) on the route, such that only the creator of the route (A) knows the destination (B).

Alas, when one node goes down along the way, the route has to be recreated, at least partly.
Alice will be informed of the crash by the predecessor of the failed node with a control message, using the onion route.
Then, she will have to select a new connected peer, and ask the predecessor of the failed node to re-establish the route using this new peer.
This is fine in scenarii where participating servers only fail sporadically.
But in our setup, where devices disconnection is a certainty, we need a more robust approach.

We select several relays at each point of the route: $R_i=\{r_{i,j}\}$ would be the \emph{set} of possibly disconnected devices that will take the $i^\text{th}$ position in our onion route.
In traditional onion routing, $r_1$ is the only one knowing the address of $r_2$, because it is written encrypted with its public key in the header.
Here, every device in the $R_i$ can read the set $R_{i+1}$ in the header, because the symmetric key $sk$ used to encrypt the set $R_{i+1}$ is known to every device in $R_i$.
The header indeed contains a version of $sk$ for each device $r_{i,j}$, encrypted with its own public key $pk_{r_{i,j}}$.

In the following equation, $S(sk,\texttt{message})$ means that $\texttt{message}$ is encrypted with the symmetric key $sk$, while $P(pk, \texttt{message})$ means that $\texttt{message}$ is encrypted using the public key $pk$.
Follows the content of the header in our version of the onion routing, considering that $R_i=\{r_{i,1}, r_{i,2}, r_{i,3}\}$, that $pk_{r_{i,j}}$ is $r_{i,j}$'s public key, and that $||$ means concatenation:

$$ P(pk_{r_{i,1}}, sk) || P(pk_{r_{i,2}}, sk) || P(pk_{r_{i,3}}, sk) || S(sk, R_{i+1}) $$

(I read that this is how PGP handles multiple recipients.)

% subsubsection onion_routes_in_face_of_churn (end)

% \begin{figure}[t]
% \centering
% \includegraphics[width=0.8\columnwidth]{figures/schema.pdf}
% \caption{\label{fig:pipeline}Pipeline of the steps undergone by Alice when he shares a file from his laptop L with Bob.}
% \end{figure}

% \Cref{fig:pipeline} depicts the process of a file exchange from Alice to Bob.
% The file $f$ is initially located on Alice's laptop $L$, which is potentially offline for now.
% Firstly, Alice and Bob need to exchange encryption keys and the key $ID_f$ 
% referencing the file on the \ac{DHT}.
% They do so \emph{out-of-band}, that is using another channel than the one 
% presented here (\eg bluetooth, email, \etc).

% Then, Alice queries the DHT for devices that will most likely be online at the same time as his laptop $L$.
% To this end, Alice uses his knowledge on his devices connection patterns $A_i$, along with the DHT's public information on devices connections.
% Once Alice and the DHT agreed upon a set of peers $(M_1, M_2, M_3)$, their address is put in the DHT at the key $ID_f$ referencing the file of interest.

% At the third step, Alice's laptop $L$ becomes available. 
% Informed of the ongoing file exchange by Alice's Device Swarm Protocol, $L$ fetches the addresses of $M_1, M_2$ and $M_3$ on the DHT using the key $ID_f$.
% These peers will serve as middle-men: $L$ sends them redundant chunks of $f$ through an onion route (thus remaining anonymous), for Bob to fetch them later.

% In the last forth step, Bob fetches the file $f$ using any of her devices.
% To do so, she uses the key $ID_f$ to learn the addresses of $M_1, M_2$ and $M_3$.
% Then, she fetches parts of $f$ from the online middle-men through an onion route.
% Given that $L$ sent redundant chunks to the middle-men, Bob will not need every middle-man to be connected to fetch the whole file.
% Once Bob has completed the download, the key $ID_f$ is flushed on the DHT.

### 2018-04-12 paper/intro.tex
finished route_success_rate plot
described my new idea

removed big'o'file

>>>>

\commentAL{The proposal:}We propose \name, a \ac{P2P} system for secret file exchange.
Using only devices owned by participants, \name creates onion routes that allow two users to exchange a file such that nobody is able to determine the content of the file, nor its source and destination.
Personal devices, unlike servers, often disconnect from the network. To permit reliable file exchange despite the unreliability of the nodes constituting the system, we use devices' knowledge on their owner's behavior to predict future (dis)connections.

### 2018-04-12 paper/user-model.tex
finished route_success_rate plot
described my new idea

removed big'o'file

We consider a single user, owning a set of $\mathcal{D}$ devices.
Time is discretized, and the devices' observations are independent: Alice can have her laptop and phone connected at the same time.
The hidden process $S=S_1,\dots,S_T$ is built from an alphabet of states $\mathcal{S}$, while the observable process $O=O_1, \dots, \O_T$ contains events from the alphabet $\mathcal{O}$. We define a multivariate HMM $\lambda_N$ as follows:

@@

\subsubsection{Predicting future behavior} % (fold)
\label{ssub:predicting_future_behavior}
\commentAL{Why we need to make these predictions is not explained yet.}
Devices want to predict whether they are likely to be connected in the near future.
This information will be publicly available, and will allow routes selection for the  files exchange between users.

@@

It is the sum of the probabilities, for each state $s$, that the user switches to $s$ and uses her device.
Recursively, predicting $i$ turns in advance follows the same logic:

@@

>>>>

We consider a single user, owning a set of $\mathcal{D}$ devices.
Time is discretized \commentAL{Discretized/synchronous time needs more explanation}. Needs, and the devices' observations are independent: Alice can have her laptop and phone connected at the same time.
The hidden process $S=S_1,\dots,S_T$ is built from an alphabet of states $\mathcal{S}$, while the observable process $O=O_1, \dots, \O_T$ contains events from the alphabet $\mathcal{O}$. We define a multivariate HMM $\lambda_N$ as follows:

@@

\subsubsection{Predicting future behavior} % (fold)
\label{ssub:predicting_future_behavior}
Devices want to predict whether they are likely to be connected in the near future.
This information will be publicly available, and will allow routes selection for the  files exchange between users.

@@

It is the sum of the probabilities, for each state $s$, that the user switches to $s$ and uses her device.
\commentAL{Following maybe not needed if we only use $t+1$:}
Recursively, predicting $i$ turns in advance follows the same logic:

@@

\commentAL{In conclusion: we want more unicorns at the Olympic Games.}

\commentAL{I don't mention prediction yet since I am not sure we get good enough results.}
